Decoding Sleep Disorders Through Self-Reported Patterns

### 2024-04-19 paper/paper.qmd
grammar fixes

Utilizing the National Health and Nutrition Examination Survey (NHANES) dataset from 2017 to March 2020, our research examines how various sleep behaviors—such as snoring frequency, daytime sleepiness, and sleep regularity—correlate with the likelihood of reporting sleep issues to healthcare providers. The primary estimand of our analysis is the association between these behaviors and the probability of an individual reporting sleep disturbances, quantified through logistic regression. This approach provides a clearer understanding of the epidemiology of sleep problems and identifies key predictive factors.

@@

The dataset underpinning this analysis is derived from the National Health and Nutrition Examination Survey (NHANES), spanning from 2017 to March 2020.
This public dataset includes responses from participants regarding their sleep patterns, incorporating 10,195 records initially.
After meticulous data cleaning, the dataset for analysis stands at 10,031 records, encapsulating variables critical to our research: the respondent's ID, usual sleep and wake times on both weekdays and weekends, total sleep duration, frequency of snoring, incidence of breathing pauses during sleep, and self-reported communication of sleep troubles to a health professional.
The dataset provides a snapshot of Americans' sleep behaviors before the disruption caused by the COVID-19 pandemic [@Prevention_2021].

The survey participants' ages range widely, reflecting the diversity of the American population.
Variables are finely tuned to capture the multifaceted nature of sleep, encompassing aspects such as duration, disruptions, and subjective experiences of daytime sleepiness.
The NHANES protocol ensures that this dataset is a robust and reliable source of information, adhering to stringent ethical standards and data collection methods, as detailed in the NHANES Analytic Guidelines.
For further information on the data cleaning specifics and validation checks, please see the supplementary material in Appendix.

@@

Our statistical exploration was conducted within the R programming environment [@r], leveraging its comprehensive ecosystem for data analysis.
We utilized the tidyverse collection of R packages [@rTidyverse] to streamline our data processing tasks.
The ggplot2 package [@rGgplot2] was instrumental in crafting insightful visualizations that articulated the intricate relationships within our data.
The dplyr package [@rDplyr] provided a syntax that facilitated the manipulation and transformation of our dataset, enabling us to prepare the data effectively for logistic regression analysis.
Data importation was efficiently handled by the readr package [@rReadr], known for its quick and user-friendly approach to reading tabular data.
Navigational simplicity within our project's directories was achieved with the here package [@R-here], which reliably managed file paths without the need for manual path setting.
The reproducibility of our research was enhanced by the knitr package [@rKnitr], which seamlessly wove R code into our report, ensuring that our findings are transparent and replicable.
For tabular data presentation, kableExtra[@R-kableExtra] offered a suite of customization options that enhanced the readability and aesthetic appeal of our tables.
The logistic regression model was developed using core functions in R, which provide robust methods for estimating the effects of various predictors on a binary outcome.

@@

## Measurement:

@@

Let $y_i$ denote the binary outcome indicating whether an individual has reported sleep troubles to a doctor, with $y_i = 1$ for reported troubles and $y_i = 0$ otherwise.
The predictors include weekday sleep duration ($x_{i1}$), weekend sleep duration ($x_{i2}$), snoring frequency ($x_{i3}$), breathing pause frequency ($x_{i4}$), and daytime sleepiness ($x_{i5}$).
The logistic regression model is formulated as follows:

@@

The priors for the intercept ($\beta_0$) and the slopes are deliberately chosen to reflect initial neutrality or skepticism about each predictor's impact on the likelihood of reporting sleep troubles. Specifically, $\beta_0$, $\beta_1$, and $\beta_2$ are assigned Normal(0, 1) priors to maintain an unbiased stance before observing the data, acknowledging that the number of sleep hours could either positively or negatively influence sleep troubles, with only a modest expectation of effect size.

For $\beta_3$, $\beta_4$, and $\beta_5$, which relate to snoring frequency, breathing pause frequency, and overly sleep frequency, the same Normal(0, 1) priors are used. This choice indicates no strong preconceived notions about the direction or magnitude of their impact, allowing the data itself to guide the inference about these predictors without the influence of strong initial assumptions.

@@

Our Bayesian logistic regression model is designed to investigate the associations between various sleep-related factors and the likelihood of reporting sleep troubles.
The priors for the coefficients are set with specific hypotheses in mind, informed by existing literature and plausible biological mechanisms:

@@

## Model Equation

@@

## Predict Sleep Disorder Through Self Reported Patterns

>>>>

Utilizing the National Health and Nutrition Examination Survey (NHANES) dataset from 2017 to March 2020 [@Prevention_2021], our research examines how various sleep behaviors—such as snoring frequency, daytime sleepiness, and sleep regularity—correlate with the likelihood of reporting sleep issues to healthcare providers. The primary estimand of our analysis is the association between these behaviors and the probability of an individual reporting sleep disturbances, quantified through logistic regression. This approach provides a clearer understanding of the epidemiology of sleep problems and identifies key predictive factors.

@@

The dataset used in this analysis is derived from the National Health and Nutrition Examination Survey (NHANES), spanning from 2017 to March 2020.
This public dataset includes responses from participants regarding their sleep patterns, incorporating 10,195 records initially.
After some data cleaning, the dataset for analysis stands at 10,031 records, encapsulating variables critical to our research: the respondent's ID, usual sleep and wake times on both weekdays and weekends, total sleep duration, frequency of snoring, incidence of breathing pauses during sleep, and self-reported communication of sleep troubles to a health professional.
The dataset provides a snapshot of Americans' sleep behaviors before the disruption caused by the COVID-19 pandemic [@Prevention_2021].

The survey participants' ages range widely, reflecting the diversity of the American population.
Variables are finely tuned to capture the multifaceted nature of sleep, including aspects such as duration, disruptions, and subjective experiences of daytime sleepiness.
The NHANES protocol ensures that this dataset is a robust and reliable source of information, assure ethical standards and data collection methods, as detailed in the NHANES Analytic Guidelines.
For further information on the data cleaning specifics and validation checks, please see the supplementary material in Appendix.

@@

Our analytical approach was completed using R [@r], utilizing the tidyverse [@rTidyverse] for data manipulation and ggplot2 [@rGgplot2] for visualization. The readr [@rReadr], here [@R-here], and arrow [@rArrow] packages were pivotal in handling data importation and directory management. Reproducible reporting was facilitated by knitr [@rKnitr] and kableExtra [@R-kableExtra] for enhancing table presentations. Bayesian logistic regression modeling was conducted with rstanarm [@rRstanarm], while model diagnostics were visualized using bayesplot [@rBayesplot]. These tools, which formed the backbone for this paper as well as the scripts, ensured efficient and reproducible analyses, culminating in a robust set of findings.

@@

\newpage
## Measurement:

@@

Let $y_i$ denote the binary outcome indicating whether an individual has reported sleep troubles to a doctor, with $y_i = 1$ for reported troubles and $y_i = 0$ otherwise.
The predictors include continous variables like weekday sleep duration ($x_{i1}$), and weekend sleep duration ($x_{i2}$); ordinal categorical variables such as snoring frequency ($x_{i3}$), breathing pause frequency ($x_{i4}$), and overly sleep frequency ($x_{i5}$).
The logistic regression model is formulated as follows:

@@

The priors for the intercept ($\beta_0$) and the slopes are deliberately chosen to reflect initial neutrality or skepticism about each predictor's impact on the likelihood of reporting sleep troubles. Specifically, $\beta_0$ for intercept, $\beta_1$, and $\beta_2$ for the continous variables about sleep hours are assigned Normal(0, 1) priors to maintain an unbiased stance before observing the data, acknowledging that the number of sleep hours could either positively or negatively influence sleep troubles, with only a modest expectation of effect size.

For $\beta_3$, $\beta_4$, and $\beta_5$, which relate to ordinal categorical variables such as snoring frequency, breathing pause frequency, and overly sleep frequency are set to have the same Normal(0, 1) priors. This choice indicates no strong preconceived notions about the direction or magnitude of their impact, allowing the data itself to guide the inference about these predictors without the influence of strong initial assumptions.

@@

Our Bayesian logistic regression model is designed to investigate the associations between various sleep-related factors and the likelihood of reporting sleep troubles.
The priors for the coefficients are set with specific hypotheses in mind, informed by knowleged gained from the data exploration section:

@@

## Logistic Regression Model for Predicting Individual Sleep Disorder Probability

@@

## Example of Predict Sleep Disorder Through Self Reported Patterns
